Learning and comparing functional connectomes between populations

### 2012-12-08 paper.tex
Progress on the parcellation part

@@

@@

Different strategies: 
dense/sparse: covering a large fraction of the
brain or not == tradeoff between functional specificity and covering the
full brain
, hard vs soft boundaries (spatial regression?).

\paragraph{Regions from atlases}

AAL \cite{tzourio-mazoyer2002a}, Harvard-oxford, sulci-based atlas.
Personal bias against the AAL, but note that it is widely used because of
SPM toolbox.

\paragraph{Defining regions from the literature}
Meta-analysis \cite{dosenbach}, DMN/TPN ROIs (Cameron, who is that guy?)

\paragraph{FMRI-based function definition}
Distinction between activation and rest.
Activation: thresholding GLM maps, or using spheres around the activation
peak.

Rest: ICA-based approaches \cite{kiviniemi2009} \cite{shirer2012}
\cite{varoquaux2011} \cite{yeo}, cite Power?
Maybe use the ICA from the Smith 2009 paper

Clustering approaches, \cite{craddock2011}, \cite{bellec}, \cite{flandin}, \cite{poline}, \cite{thirion}

@@

\caption{
Different regions
}
\end{figure}

@@

% Disclaimer: focus on correlation/second order statistics?
Mutual information? Entropy correlation? MI as measure of integration (Marralec)

@@

>>>>

Define what we call a functional connectome, and the term \emph{graph},
as well as its relation with an adjacency matrix.

Define a functional network

@@

Here we discuss strategies to infer connectomes from functional brain
imaging data. We start with the choice of nodes \emph{i.e.} regions,
followed by signal extraction, and the estimation of graphs.

@@

The choice of regions of interests (ROIs) that defines the nodes of the
graphs can be very important both in the estimation of connectomes and
for group comparison \cite{wang2009}. Unsurprisingly, simulations have
shown that extracting signal from ROIs that did not match functional
units would lead to incorrect graph extraction \cite{smith2011}.
%
Different strategies coexist. While dense parcellations approaches cover
a large fraction of the brain \cite{achard2006, varoquaux2010c,
wang2009}, this coverage can be traded off to focus on some specific
regions, in favor of increased functional specificity and thus better
differentiation across networks \cite{greicius2003, fair2009,
varoquaux2010b, lewis2009, fransson2008, shirer2012}. In addition, while
ROIs and most often defined as a hard selection of voxels, it is also
possible to use a \emph{soft} definition, attributing weights as with
probabilistic atlases, or spatial maps of functional networks extracted
from techniques such as independent component analysis (ICA).

\paragraph{Regions from atlases}
%
Atlases can be used to define full-brain parcellations. The most popular
choice is probably the AAL atlas \cite{tzourio-mazoyer2002a}, that
benefits from an SPM toolbox. However, it suffers from major
shortcomings; namely that \emph{i)} it was defined on a single subject
and thus does does not reflect inter-subject variability and \emph{ii)}
focuses on labeling large anatomical structures and does not match
functional layout --for instance only two regions describe the medial
part of the frontal lobe. Multi-subject probabilistic altases such as the
Harvard-Oxford atlas distributed with FSL \cite{smith2004} or the
sulci-based structural atlas used in \cite{varoquaux2010c} mitigate the
first problem, and the high number of regions defined using sulci also
somewhat circumvent the second problem (see
fig.\,\ref{fig:parcellations}).

\paragraph{Defining regions from the literature}
%
Regions can be defined from previous studies, informally or with
systematic meta-analysis. This strategy is used to define the main
resting-state networks, such as the default mode network, but may also be
useful to study connectivity in task-specific networks
\cite{grillon2012}. The common practice is to place balls of a given
radius, 5 or 10\,mm, centered at the coordinates of interest. Given that
functional networks are tightly interleaved in some parts of the cortex,
such as the parietal lobe, care must be taken not to define too many
regions that would overlap and lead to mixing of the signal.

\paragraph{FMRI-based function definition}
%
Defining regions directly from the fMRI signal brings many benefits.
First, it can capture subject-specific functional information. Second, it
also adapts to the signal at hand and it's limitations, such as
EPI-specific warps or vascular and movement artifacts that are isolated
in ICA-like approaches. The simplest approach to define task-specific
regions is to use activation maps derived from standard GLM-based
analysis in a task-driven study (see for instance \cite{poldrack2011}).
Regions are extracted by thresholding the maps, or using balls around the
activation peaks. For resting-state studies, unsupervised multivariate
analysis techniques are necessary. Clustering approaches extract
full-brain parcellations \cite{craddock2011, bellec2010, yeo2011,
thirion2006}, and have been shown to extract well-known functional
structures from rest data \cite{yeo2011}. Alternatively, decompositions
methods, such as ICA \cite{beckmann2004}, can unmixing linear
combinations of multiple effects and separate out multiple spatial maps
capturing functional networks or confounding effects. At high model
order, ICA maps define a functional parcellation \cite{kiviniemi2009}.
Extracting regions from these maps requires additional effort as they can
display fragmented spatial features and structured background noise, but
incorporating sparsity and spatial constraints in the decomposition
techniques leads to contrasted maps that outline many different
structures \cite{varoquaux2012} (see fig.\,\ref{fig:parcellations}).

@@

\caption{
Different full-brain parcellation: the AAL atlas
\cite{tzourio-mazoyer2002a}, regions extracted by Ncuts
\cite{craddock2011}, the Harvard-Oxford atlas, the sucli atlas used in
\cite{varoquaux2010c}, the resting-state networks extracted in
\cite{smith2009} by ICA, and in \cite{varoquaux2011} by sparse dictionary
learning.
\label{fig:parcellations}
}
\end{figure}

@@

% Disclaimer: focus on correlation/second order statistics?

@@

Mutual information? Entropy correlation? MI as measure of integration (Marralec)
